On the problem of repeated supervised learning

### 2023-04-12 paper/M1P_Vepricov/main.tex
cross-check

\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}         % Can be removed after putting your text content
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsthm}
\usepackage{amssymb}
\newtheorem{theorem}{Theorem}

@@

@@

        Now write out a Herder's inequality. For $q$ such that $\frac{1}{p} + \frac{1}{q} = 1$ is fulfilled that

@@

    To begin with, note that the result of Theorem \ref{ineq_q} does not depend in any way on whether $\text{D}$ converts $\mathbf{R}$ to $\mathbf{R}$, it is a consequence of Herder's inequality.

@@

    \subsection{Experiment Design} \label{design}
        We consider such a formulation of the problem of repeated supervised learning: there is an original data $\textbf{X}$ and a vector of target variables $\mathbf{y}$. Initially, at round $r = 0$, we select $30\%$ of the original dataset on which our model $H(\mathbf{x}, \theta^0, 0)$ is trained with $80\%$ train size.

@@

    \subsection{Experiment results} \label{res}

@@

>>>>

\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}         % Can be removed after putting your text content
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsthm}
\usepackage{amssymb}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}

@@

    How we provide several Lemmas based on Theorem \ref{delta}. In In these lemmas we assume that the operator $\text{D}$ has the form

    \begin{equation} \label{cool_D}
        \text{D}^t(f_0)(x) = \psi(t) \cdot f_0(\psi(t) \cdot x) ~\text{ and }~ \psi(t) \to +\infty
    \end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{lemma}[Conditions on $\{\text{D}^t\}_{t=0}^{+\infty}$ to be a semigroup] \label{semigroup}

        If $\text{D}$ has the form \eqref{cool_D}, then $\{\text{D}^t\}_{t=0}^{+\infty}$ is a semigroup, i.e. $(\text{D}^{\tau} \circ \text{D}^{\kappa})(f) = \text{D}^{\tau + \kappa}(f) ~\forall \tau, \kappa \in \mathbb{N}$, if and only if

        \begin{equation} \label{cond_semigroup}
            \psi(\tau + \kappa) = \psi(\tau) \cdot \psi(\kappa) ~\forall \tau, \kappa \in \mathbb{N}
        \end{equation}

    \end{lemma}

    \begin{proof}
        \begin{equation*}
            (\text{D}^{\tau} \circ \text{D}^{\kappa})(f)(x) = \text{D}^{\tau}(\psi(\kappa) \cdot f(\psi(\kappa) \cdot x)) = \psi(\tau)\psi(\kappa) \cdot f(\psi(\tau)\psi(\kappa) \cdot x)
        \end{equation*}

        \begin{equation*}
            \text{D}^{\tau + \kappa}(f)(x) = \psi(\tau + \kappa) \cdot f(\psi(\tau + \kappa) \cdot x)
        \end{equation*}

        So, $\{\text{D}^t\}_{t=0}^{+\infty}$ is a semigroup $\Leftrightarrow$ $\psi(\tau + \kappa) = \psi(\tau) \cdot \psi(\kappa) ~\forall \tau, \kappa \in \mathbb{N}$
    \end{proof}

    \begin{lemma}[Decreasing moments] \label{moments}
        If $\text{D}$ has the form \eqref{cool_D}, then all $k$-th moments of random variable $\|\mathbf{y} - \mathbf{y_{\text{pred}}}\|$ (if they exist) are decreasing with speed $\psi(t)^{-k}$, i.e. $\nu_k^t = \psi(t)^{-k} \nu_k^0$, where $\nu_k^t$ is a $k$-th moment on a step $t$.

        If $~\exists q \in [1; +\infty]$ such $\{\nu_k^0\}_{k=1}^{+\infty} \in l_q$, then $\{\nu_k^t\}_{k=1}^{+\infty} \in l_1$ and $\{\nu_k^t\}_{k=1}^{+\infty} \underset{t \to \infty}{\overset{l_1}{\longrightarrow}} 0$
    \end{lemma}

    \begin{proof}
        Let's first prove first term. By the definition of $k$-moment we have

        \begin{equation*}
            \nu_k^t = \int\limits_{-\infty}^{+\infty} x^k \psi(t) f(\psi(t) x) dx
        \end{equation*}

        If we make variable substitution $y = \psi(t) x$, when we have

        \begin{equation*}
            \nu_k^t = \int\limits_{-\infty}^{+\infty} \dfrac{y^k}{\psi(t)^k} f(y) dy = \psi(t)^{-k} \nu_k^0
        \end{equation*}

        So the first term is proved. Consider the second term.

        \begin{equation*}
            \|\{\nu_k^t\}_{k=1}^{+\infty}\|_1 = \|\{\psi(t)^{-k} \nu_k^0\}_{k=1}^{+\infty}\|_1 \leq \|\{\psi(t)^{-k}\}_{k=1}^{+\infty}\|_p \cdot \|\{\nu_k^0\}_{k=1}^{+\infty}\|_q
        \end{equation*}

        The second step follows from Helder's inequality.

        How let's calculate $\|\{\psi(t)^{-k}\}_{k=1}^{+\infty}\|_p$ for $p \in [1; +\infty)$:

        \begin{equation*}
            \|\{\psi(t)^{-k}\}_{k=1}^{+\infty}\|_p^p = \sum\limits_{k=1}^{+\infty}\psi(t)^{-kp} = \dfrac{\psi(t)^{-p}}{1 - \psi(t)^{-p}} = \dfrac{1}{\psi(t)^p - 1}
        \end{equation*}

        The first equality is true only if $\psi(t) > 1$ and the second step follows from sum of infinitely decreasing geometric progression.

        \begin{equation*}
            \|\{\psi(t)^{-k}\}_{k=1}^{+\infty}\|_p = \left( \dfrac{1}{\psi(t)^p - 1} \right)^{1/p} \underset{t \to +\infty}{\longrightarrow} 0 ~~~\forall p \in [1; +\infty)
        \end{equation*}

        If $p = +\infty$:

        \begin{equation*}
            \|\{\psi(t)^{-k}\}_{k=1}^{+\infty}\|_{\infty} = \left[ \text{ if } \psi(t) > 1 \right] = \psi(t)^{-1} \underset{t \to +\infty}{\longrightarrow} 0
        \end{equation*}

        So, we have

        \begin{equation*}
            \|\{\nu_k^t\}_{k=1}^{+\infty}\|_1 \leq \|\{\psi(t)^{-k}\}_{k=1}^{+\infty}\|_p \cdot \|\{\nu_k^0\}_{k=1}^{+\infty}\|_q \underset{t \to +\infty}{\longrightarrow} 0
        \end{equation*}

        Because $\|\{\nu_k^0\}_{k=1}^{+\infty}\|_q < +\infty$ as a condition of Lemma.

%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \textbf{Discussion of Lemmas \ref{semigroup} and \ref{moments}}

    If our system is autonomous, i.e. it does not depend on time, then the operators $\{\text{D}^t\}_{t=0}^{+\infty}$ should form a semigroup, since their application should not depend on the number of step $t$. The condition \eqref{cond_semigroup} means that the function $\psi(t)$ is a power function.

    The Lemma \ref{moments} on moments is interesting from the practical point of view as a condition which is relatively easy to check in the experiment, which we will do in the Section \ref{Experiments}.

@@

        Now write out a Helder's inequality. For $q$ such that $\frac{1}{p} + \frac{1}{q} = 1$ is fulfilled that

@@

    To begin with, note that the result of Theorem \ref{ineq_q} does not depend in any way on whether $\text{D}$ converts $\mathbf{R}$ to $\mathbf{R}$, it is a consequence of Helder's inequality.

@@

    \subsection{Experiment Design} \label{design}
        We consider \footnotemark such a formulation of the problem of repeated supervised learning: there is an original data $\textbf{X}$ and a vector of target variables $\mathbf{y}$. Initially, at round $r = 0$, we select $30\%$ of the original dataset on which our model $H(\mathbf{x}, \theta^0, 0)$ is trained with $80\%$ train size.

        \footnotetext{All experiments you can see on our \href{https://github.com/intsystems/2023-Project-119}{GitHub}}

@@

    \subsection{First experiment: analysis of deviation} \label{exp_1}

@@

    \subsection{Second Experiment: analysis of $p$-value} \label{exp_2}
        In this experiment we check our data for belonging to a normal distribution by counting $p$-value from normal test and we also count the mean and variance of the data $\mathbf{y} - \mathbf{y_{\text{pred}}}$ at each step $t$. We take \textit{usage} and \textit{adherence} \ref{ex_set} based on the previous experiment \ref{exp_1}: $1.0$ and $0.0$, $0.6$ and $0.8$, $0.1$ and $0.9$. The results of this experiment are shown in the Figure \ref{fig_exp_2}.

        \begin{figure}[h!]
            \centering
            \includegraphics[width=0.49\linewidth]{pictures/mdp_0.1_0.9.png}~~
            ~~\includegraphics[width=0.49\linewidth]{pictures/mdp_0.6_0.8.png}

            \includegraphics[width=0.49\linewidth]{pictures/mdp_1_0.png}

            \caption{Results of experiment \ref{exp_2} \textit{usage} and \textit{adherence} = $0.1$ and $0.9$ (left), $0.6$ and $0.8$ (right), $1.0$ and $0.0$ (bottom).}
            \label{fig_exp_2}
        \end{figure}

        As we can see for small \textit{usage} the deviation of out data does not decrease, but $p$-value for all \textit{usage} and \textit{adherence} is bigger than $0.05$, so we can assume that our data are normally distributed at each step $t$.

    \subsection{Third experiment: limit to delta function} \label{exp_3}
        In this experiment we test the conditions from Theorem \ref{delta}, i.e. we measure $f_t(0)$ and $\int\limits_{-\kappa}^{\kappa}f_t(x)dx = \hat{F}_t(\kappa) - \hat{F}_t(-\kappa)$, where $\kappa$ is sufficiently small. So if $f_t(0) \to +\infty$ and $1 \in I_{\varepsilon}(\hat{F}_t(\kappa) - \hat{F}_t(-\kappa))$, where $I_{\varepsilon}$ is confidence interval from \eqref{inter}, then we can say that operator $\widetilde{\text{D}}$ converge empirical distribution functions $f_t(x)$ of $\|\mathbf{y} - \mathbf{y_{\text{pred}}}\|$ to delta-function. The results of this experiment are shown in the Figure \ref{fig_exp_3}.

        \begin{figure}[h!]
            \centering
            \includegraphics[width=0.49\linewidth]{pictures/f_t0_0.1_0.9.png}~~
            ~~\includegraphics[width=0.49\linewidth]{pictures/f_t0_0.6_0.8.png}

            \includegraphics[width=0.49\linewidth]{pictures/f_t0_1_0.png}

            \caption{Results of experiment \ref{exp_3} \textit{usage} and \textit{adherence} = $0.1$ and $0.9$ (left), $0.6$ and $0.8$ (right), $1.0$ and $0.0$ (bottom).}
            \label{fig_exp_3}
        \end{figure}

        As we can see, for small values of \textit{usage} and large values of \textit{adherence}, the operator $\widetilde{\text{D}}$ does not translate the distribution density function of $\|\mathbf{y} - \mathbf{y_{\text{pred}}}\|$ to delta-function. When \textit{usage} = 1 and \textit{adherence} = 0 we can observe a tendency towards the delta function at small $t$, and then exit to a plateau. From results of this experiment we can reject the condition $\psi(t) \underset{t \to +\infty}{\longrightarrow} +\infty$ and consider $\psi(t)$ as an arbitrary function that satisfies the inequality \eqref{psi_and_g}, but then there will be no limit to the delta function when.

    \subsection{Fourth experiment: semigroup check} \label{exp_4}

        In this experiment we test our system for autonomy, that is, we test the condition \eqref{cond_semigroup} in the Lemma \ref{semigroup}. As noted in the Discussion of Lemma \ref{semigroup}, in order for the $\psi(t)$ function to satisfy the condition \eqref{cond_semigroup}, it is necessary and sufficient that $\psi(t)$ should be a power function. Therefore, in this experiment we plot $\ln(\psi(t))$, and if we get a straight line, then the system is autonomous, and if not, then no. The $\psi(t)$ function is taken from the Experiment \ref{exp_3}. The results of this experiment are shown in the Figure \ref{fig_exp_4}.

        \begin{figure}[h!]
            \centering
            \includegraphics[width=0.49\linewidth]{pictures/semigroup_0.1_0.9.png}~~
            ~~\includegraphics[width=0.49\linewidth]{pictures/semigroup_0.6_0.8.png}

            \includegraphics[width=0.49\linewidth]{pictures/semigroup_1_0.png}

            \caption{Results of experiment \ref{exp_4} \textit{usage} and \textit{adherence} = $0.1$ and $0.9$ (left), $0.6$ and $0.8$ (right), $1.0$ and $0.0$ (bottom).}
            \label{fig_exp_4}
        \end{figure}

        As we can see, our system is not autonomous, since none of the graphs is a straight line.

    \subsection{Fifth experiment: decreasing in moments} \label{exp_5}
        In this experiment we test the results from Lemma \ref{moments}. Based on the Experiment \ref{exp_2} we will approximate our distribution by a normal distribution. Based on Wick's theorem, we can explicitly find the values of moments for the normal distribution if its mathematical expectation is 0:

        \begin{equation} \label{normal_moments}
            \nu_k = \left\{\begin{array}{ll}
                 0, \text{ if } k = 2n + 1\\
                 \\
                 \sigma^k \cdot (k - 1)!!, \text{ if } k = 2n
            \end{array} \right.
        \end{equation}

        The results of this experiment are shown in the Figure \ref{fig_exp_5}.

        \begin{figure}[h!]
            \centering
            \includegraphics[width=0.49\linewidth]{pictures/moments_0.1_0.9.png}~~
            ~~\includegraphics[width=0.49\linewidth]{pictures/moments_0.6_0.8.png}

            \includegraphics[width=0.49\linewidth]{pictures/moments_1_0.png}

            \caption{Results of experiment \ref{exp_5} \textit{usage} and \textit{adherence} = $0.1$ and $0.9$ (left), $0.6$ and $0.8$ (right), $1.0$ and $0.0$ (bottom).}
            \label{fig_exp_5}
        \end{figure}

        The moments decrease to zero only if \textit{usage} and \textit{adherence} = $1$ and $0$, because only with this parameters operator $\text{D}$ transforms density distribution functions of $\|\mathbf{y} - \mathbf{y_{\text{pred}}}\|$ to delta function.
